State of the Practice for Medical Imaging Software

### 2022-11-02 StateOfPractice/Papers/MedImageSoft_SOP-arXiv/MedImageSoft_SOP.tex
Final draft for submission, next step to anonymize

%\usepackage{tikz}
%\usetikzlibrary{mindmap}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{arydshln}
\usepackage{paralist}
\usepackage{booktabs}

@@

We present the state of the practice for Medical Imaging (MI) software. We
selected 29 medical imaging projects from 48 candidates, assessed 10 software
qualities (installability, correctness/ verifiability, reliability, robustness,
usability, maintainability, reusability, understandability,
visibility/transparency and reproducibility) by answering 108 questions for each
software project, and interviewed 8 of the 29 development teams. Based on the
quantitative data for the first 9 qualities, we ranked the MI software with the
Analytic Hierarchy Process (AHP). The four top ranked software products are:
\textit{3D Slicer}, \textit{ImageJ}, \textit{Fiji}, and \textit{OHIF Viewer}.
Our ranking is mostly consistent with the community's ranking, with four of our
top five projects also appearing in the top five of a list ordered by
stars-per-year. \textit{DICOM Viewer} and \textit{dmv} are exceptions; we ranked
them lower than their estimated popularity because of difficulty with their
installation.  Generally, MI software is in a healthy state.  The documentation
artifacts created, and development tools employed, typically match the
recommendations given in research software development guidelines. A few places
in the current state of the practice that deviate from the recommendations
include some underemphasized artifacts (like a developer's code of conduct, code
style guidelines, product roadmaps, and Application Program Interface (API)
documentation) and ... Exceptions include ... tools With respect to software
artifacts, MI software generally Comparison for tool usage. Comparison for
process. From interviewing the developers, we identified five pain points and
two qualities of potential concern:: i) lack of resources; ii) difficulty
balancing between compatibility, maintainability, performance, and security;
and, iii) lack of access to real-world datasets for testing. For future MI
software projects, we propose adopting test-driven development, using Continuous
Integration and Continuous Delivery (CI/CD), using git and GitHub, maintaining
good documentation, supporting third-party plugins or extensions, considering
web application solutions, and improving collaboration between different MI
software projects.
\wss{Update after the paper has been revised.}

@@

\begin{table}[!ht]
\begin{center}
\begin{tabular}{ p{2.5cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1.2cm}p{1cm}p{0.8cm} }
\toprule
~ \ & \cite{USGS2019} & \cite{TobiasEtAl2018} & \cite{BrettEtAl2021} &
\cite{WilsonEtAl2016} & \cite{SmithAndRoscoe2018} & \cite{HerouxEtAl2008} &
\cite{ThielEtAl2020} & \cite{vanGompelEtAl2016} & \cite{OrvizEtAl2017} & MI\\
\midrule
LICENSE & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & &
\checkmark & \checkmark & \checkmark & C\\
README &  & \checkmark & \checkmark & \checkmark & \checkmark & & \checkmark &
\checkmark & \checkmark & C\\
CONTRIBUTING &  & \checkmark & \checkmark & \checkmark & \checkmark & &
\checkmark & \checkmark & \checkmark & R\\
CITATION &  &  &  & \checkmark & & & & \checkmark & \checkmark & U\\
CHANGELOG &  & \checkmark &  & \checkmark & \checkmark & & \checkmark &  &  & U\\
INSTALL &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark & U\\
\midrule
Uninstall &  &  &  &  & & & & \checkmark & &  \\
Dependency List &  &  & \checkmark & & \checkmark & & & \checkmark &  & R\\
Authors &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & U\\
Code of Conduct &  &  &  &  & & & \checkmark & & & R\\
Acknowledgements &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & U\\
Code Style Guide &  & \checkmark &  &  & & & \checkmark & \checkmark & \checkmark & R\\
Release Info. &  & \checkmark &  &  & & \checkmark & \checkmark & & & C\\
Prod.\ Roadmap &  &  &  &  & & \checkmark & \checkmark & \checkmark & & R\\
\midrule
Getting started &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark & R\\
User manual &  &  & \checkmark &  & & & \checkmark & & & C\\
Tutorials &  &  &  &  & & & \checkmark & & & U\\
FAQ &  &  &  &  & & & \checkmark & \checkmark & \checkmark & U\\
\midrule
Issue Track &  & \checkmark & \checkmark & & \checkmark & \checkmark &
\checkmark & & \checkmark & C\\
Version Control &  & \checkmark & \checkmark & \checkmark & \checkmark &
\checkmark & \checkmark & \checkmark & \checkmark & C\\ 
Build Scripts &  & \checkmark &  & \checkmark & \checkmark & \checkmark &
\checkmark & & \checkmark & U\\
\midrule
Requirements &  & \checkmark &  &  & & \checkmark &  &  & \checkmark & R\\
Design Doc.\ &  & \checkmark  & \checkmark &  & \checkmark & & \checkmark &
\checkmark& \checkmark & R\\
API Doc. &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark & R\\
Test Plan &  & \checkmark &  &  & & \checkmark & & & &  \\
Test Cases & \checkmark & \checkmark & \checkmark &  & \checkmark & \checkmark &
\checkmark & \checkmark & \checkmark & U\\
\bottomrule
\end{tabular}
\caption{Comparison of Recommended Artifacts in Software Development Guidelines
to Artifacts in MI Projects (C for Common, U for Uncommon and R for Rare)}
\label{Tbl_Guidelines}
\end{center}
\end{table}

@@

Although the MI community shows examples of 88\% (23 of 26) of the practices we
found in research software guidelines (Table~\ref{Tbl_Guidelines}), we did not
observe three recommended artifacts: i) Uninstall, ii) Test plans, and iii)
Requirements.  Uninstall is likely an omission caused by the focus on installing
software. Given the storage capacity of current hardware, developers and users
are not generally concerned with uninstall.  Moreover, as mentioned above,
uninstall is not particularly emphasized in existing recommendations.  We did
not observe test plans for MI software, but that doesn't mean plans weren't
created; it means that the plans are not under version control.  Test plans
would have to at least be implicitly created, since we observed test cases with
reasonable frequency for MI software (test cases are categorized as uncommon).

MI software is like other research software in its neglect of requirements
documentation.  Although requirements documentation is recommended by some
\citep{TobiasEtAl2018, HerouxEtAl2008, SmithAndKoothoor2016}, in practice
research software developers often do not produce a proper requirements
specification \citep{HeatonAndCarver2015}. \citet{SandersAndKelly2008}
interviewed 16 scientists from 10 disciplines and found that none of the
scientists created requirements specifications, unless regulations in their
field mandated such a document. \citet{Nguyen-HoanEtAl2010} showed requirements
are the least commonly produced type of documentation for research software in
general. When looking at the pain points for research software developers,
\citet{WieseEtAl2019} found that software requirements and management is the
software engineering discipline that most hurts scientific developers,
accounting for 23\% of the technical problems reported by study participants.
The lack of support for requirements is likely due to the perception that
up-front requirements are impossible for research software
\citep{CarverEtAl2007, SegalAndMorris2008}, but if we drop the insistence on
``up-front'' requirements, allowing instead for the requirements to be written
iteratively and incrementally, requirements are feasible \citep{Smith2016}.

Table~\ref{Tbl_Guidelines} shows several artifacts that are rarely observed in
practice.  A theme among these rare artifacts is that many of them are related
to developers more than users.  For instance, the developer-focused, rare, and
in some cases non-observed, artifacts include the following:

@@

\item \textbf{Requirements Documentation} records the functionalities, expected
performance, goals, context, design constraints, external interfaces and other
quality attributes of the software \citep{IEEE1998}.  Templates, which provide
structure, guidelines and rules, are often used for writing requirements
documents. Example templates include \citet{ESA1991}, \citet{IEEE1998},
\citet{NASA1989}, and \citet{RobertsonAndRobertson1999Vol}. There is no
universally accepted template. \citet{SmithEtAl2007} provides a template
tailored to research software.

@@

\item \textbf{A Test Plan} describes the scope, approach, resources, and the
schedule of planned test activities \citep{VanVliet2000}.  The plan should cover
details such as the makeup of the testing team, automated testing tools and
technology to employ, the testing process, system tests, integration tests and
unit tests.  A test plan can also include the specific plans for peer code (and
documentation) review, as discussed further in Section~\ref{Sec_PeerReview}.

@@

As mentioned in Section~\ref{sec_result_correctness_verifiability}, we
identified five projects using CI/CD tools (about 17\% of the assessed
projects). We found which projects used CI/CD by examining the documentation and
source code of all projects. The count of CI/CD usage may actually be higher,
since traces of CI/CD usage may not always appear in a repository.  This was the
case for a study of LBM software, where interviews with developers showed that
more projects used CI/CD than was evident from repository artifacts alone
\citep{Michalski2021}.  The 17\% utilization for MI software contrasts with the
high frequency with which research software development guidelines recommend
continuous integration \citep{BrettEtAl2021, Brown2015, ThielEtAl2020,
Zadka2018, vanGompelEtAl2016}. Although there is currently little data available
on CI/CD utilization for research software, our impression is that CI/CD is not
yet common practice, despite its recommendation.  This is certainly the case for
LBM software, where usage numbers are similar to MI software, with only 12.5\%
of 24 LBM packages showing evidence of CI/CD in their repositories
\citep{Michalski2021}.  The survey of \citet{CarverEtAl2022} suggests higher use
of CI/CD with 54\% (54/100) of respondents indicating that they use it. However,
that survey measures something different from the current one by surveying
practitioners, rather than assessing projects.

@@

\begin{itemize}
	\item Elimination of headaches associated with a separate integration phase
	\citep{Fowler2006}, \citep[p.\ 20]{HumbleAndFarley2010}. If developers
	postpone integration, integration problems are inevitable.  Continuous
	integration means that problems are immediately obvious and the source of
	the problem can be isolated to the small increment that was just committed.
	\item Detection and removal of bugs \citep{Fowler2006} via
	automated testing.  To improve productivity, defects are best discovered and
	fixed at the point where they are introduced \citep[p.\
	23]{HumbleAndFarley2010}.  Code is not the only source of errors; they are
	also found in the files and scripts related to configuration management
	\citep[p.\ 18]{HumbleAndFarley2010}.
	\item Everyone is always working on a stable base, since the rejection of
	inadequate commits means that the main branch will always be working.  A
	stable base will always pass all tests.  If the CI system uses generators
	and linters, it will also have current documentation and coding standard
	compliant code.  A stable base improves developer productivity, allowing
	them to focus on coding, testing, and documentation.
\end{itemize}

@@

\begin{itemize}
	\item A version control system \citep{Fowler2006}. To be effective, all
	files should be under version control, not just code files.  Anything that
	is needed to build, install and run the software should be under version
	control, including configuration files, build scripts, test harnesses, and
	operating system configurations \citep[p. 19]{HumbleAndFarley2010}.
	Fortunately for the MI, as shown in Section~\ref{sec_score_maintainability}
	all our measured projects use version control.
	\item A fully automated build system \citep{Fowler2006}.  As \citet[p.\
	5]{HumbleAndFarley2010} point out, deploying software manually is an
	anti-pattern.  For MI software, Table~\ref{artifactspresent} shows 18 of 29
	packages (62\%) were observed to include build scripts.  Projects without a
	build system will need to add one to pursue using CI.
	\item An automated test system \citep{Fowler2006}. Building quality software
	involves creating automated tests at the unit, component, and acceptance
	test level, and executing these tests whenever someone makes a change to the
	code, its configuration, the environment, or the software stack that it runs
	on \citep[p.\ 83]{HumbleAndFarley2010}. As Table~\ref{artifactspresent}
	shows, test cases are in the uncommon category for MI software artifacts,
	which means that some MI projects will need to increase their testing
	automation if they wish to pursue CI.
	\item An automated system for other tasks, such as code checking,
	documentation building and web-site updating.  These other tasks are not
	essential to CI, but they can be incorporated to improve the quality
	of the code and the communication between developers and users. For
	instance, a static analysis (possibly via linters) of the code may find poor
	programming practice or lack of adherence to adopted coding standards.
	\item An integration build system to pull everything together.  Every time
	there is a check-in (for instance a pull request), the integration server
	automatically checks out the sources onto the integration machine, starts a
	build, runs tests, and informs the committer of the results. 
\end{itemize}

@@

@@

\item \textbf{Modern technologies may improve frontend performance.} Web
applications with only a frontend usually perform worse than native
applications. However, new technologies may ease this difference. For example,
some JavaScript libraries can help the frontend harness the power of the
computer's GPU and accelerate graphical computing. In addition, there are new
frameworks helping developers with cross-platform compatibility. For example,
the \href{https://flutter.dev/}{Flutter} project enables support for web,
mobile, and desktop OS with one codebase.  Other options include
\href{https://vuejs.org/} {Vue}, \href{https://angular.io/} {Angular} and
\href{https://reactjs.org/} {React} and \href{https://elm-lang.org/}{Elm}.

@@

A linter is a tool that statically analyzes code to find programming errors,
suspicious constructs, and stylistic inconsistencies \citep{Wikipedia2022_Lint}.
Linters can be used as an ad hoc check for code files, but they really come into
their own when used as part of a continuous integration system, as discussed in
Section~\ref{Sec_ContinuousIntegration}. Almost none of the research software
guidelines that we consulted, summarized in Section~\ref{Sec_CompareArtifacts},
mention linters.  The one exception is \citet{ThielEtAl2020}.  Despite the lack
of mention in the guidelines, we believe that linters have the potential to
improve code quality at a relatively low cost.

@@

\citet{Fagan1976} began work on rigorous review via code inspection.  Elements
of a typical inspection include reviewing the code against a checklist (checking
the consistency of variable names, look for terminating loops, etc.), performing
specific review tasks (such as summarizing the code's purpose, cross-referencing
the code to the technical manual, creating a data dictionary for a given module,
etc.) Rigorous inspection finds 60-65\% of latent defects on average, and often
tops 85\% in defect removal efficiency \citep{Jones2008}. The success rate of
code inspection is generally higher than most forms of testing, which average
between 30 and 35\% for defect removal efficiency \citep{EbertAndJones2009,
Jones2008}. For research software, \citep{KellyAndShepard2000} show a task based
inspection approach can be effective. Task based inspection is an ideal fit with
an issue tracking system, like GitHub.  The review tasks can be issues, so that
they can be easily assigned, monitored and recorded. Potential issues include
assigning junior developers to test getting-started tutorial and installation
instructions.

As indicated in Section~\ref{Sec_CompareMethodologies} some MI projects use
modern code review, via issue tracking and the use of GitHub.  Those MI projects
not incorporating modern code review would likely benefit by adopting it.
Although a rigorous code inspection is likely not worth the required resources,
for critical parts of the code, developers may want to adopt a more rigorous
approach. For instance, developers may drop the modern requirement for
asynchronous review and instead occasionally use synchronous review to help
uncover errors and disseminate best practices throughout the team.  For
instance, teams could periodically meet, either in-person or virtually, and have
junior members walk through their code.  In-person reviews will likely help
realize the benefits of modern code review noticed by
\citet{BirdAndBacchelli2013}: defect detection, knowledge transfer, increased
team awareness, and creation of alternative solutions to problems.

Due to improving code quality and increasing knowledge transfer, peer review
addresses the same pain points and qualities as linters
(Section~\ref{Sec_Linters}): \ppref{P_LackDevTime}, \ppref{P_TechnologyHurdles},
and \qref{Q_Maintainability}. Peer review can potentially find misunderstandings
in of how the code implements the required theory, which will improve the
software's correctness (\ppref{P_Correctness}). The benefits of peer review for
addressing pain points can be increased by extending the review from just code,
to also reviewing all software artifacts, including documentation, build
scripts, test cases and the development process itself.

@@

In our ``state of the practice'' assessment exercise for LBM software
\citep{SmithEtAl2022}, we noticed that LBM developers implicitly used
modularization based on the principle of design for change to improve
maintainability (\qref{Q_Maintainability}).  We recommend that MI developers use
the same principle for their modularizations.  Although the advice to modularize
research software to handle complexity is common \citep{WilsonEtAl2014,
StewartEtAl2017, Storer2017}, specific guidelines on how to divide the software
into modules is less prevalent.  Not every decomposition is a good design for
supporting change, as shown by \citet{Parnas1972a}.  A design with low cohesion
and high coupling \citep[p.\ 48]{GhezziEtAl2003} will make change difficult.
Especially in research software, where change is inevitable, designers need to
produce a modularization that supports change. \citep{JungEtAl2022} points out
that ocean modelling software is currently feeling the pain of not emphasizing
modularization in legacy code.

@@

As the LBM examples above illustrate, developers can accomplish design for
change by first identifying likely changes, either implicitly or explicitly, and
second by hiding each likely change behind a well-defined module interface.
This approach mirrors the recommendations from \citet{Parnas1972a}.
\citet{SmithEtAl2022} (under Section ``Documentation as Part of the Development
Process'') lists ideas for how to document the design, including the likely
changes, so that they are more visible to others.

@@

\citet{RunesonAndHost2009} defines construct validity as when the adopted
metrics represent what they are intended to measure. Our construct threats are
often related to how we assume our measurements influences the various software
qualities, as summarized in Section~\ref{sec_grading_software}. Specifically,
our construct validity related threats include the following:

>>>>

%\usepackage{tikz}
%\usetikzlibrary{mindmap}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{arydshln}
\usepackage{paralist}
\usepackage{booktabs}
\usepackage{pbox}

@@

We present the state of the practice for Medical Imaging (MI) software. We
selected 29 medical imaging projects from 48 candidates, assessed 10 software
qualities (installability, correctness/ verifiability, reliability, robustness,
usability, maintainability, reusability, understandability,
visibility/transparency and reproducibility) by answering 108 questions for each
software project, and interviewed 8 of the 29 development teams. Based on the
quantitative data for the first 9 qualities, we ranked the MI software with the
Analytic Hierarchy Process (AHP). The four top ranked software products are:
\textit{3D Slicer}, \textit{ImageJ}, \textit{Fiji}, and \textit{OHIF Viewer}.
Our ranking is mostly consistent with the community's ranking, with four of our
top five projects also appearing in the top five of a list ordered by
stars-per-year. Generally, MI software is in a healthy state as shown by the
following: in the repositories we observed 88\% of the documentation artifacts
recommended by research software development guidelines, 100\% of MI projects
use version control tools, and developers appear to use the common quasi-agile
research software development process. However, the current state of the
practice deviates from the existing guidelines because of the rarity of some
recommended artifacts (like test plans, requirements specification, code of
conduct, code style guidelines, product roadmaps, and Application Program
Interface (API) documentation), low usage of continuous integration (17\% of the
projects), low use of unit testing (about 50\% of projects), and room for
improvement with documentation (six of nine developers felt their documentation
wasn't clear enough). From interviewing the developers, we identified five pain
points and two qualities of potential concern: lack of development time, lack of
funding, technology hurdles, ensuring correctness, usability, maintainability,
and reproducibility. The interviewees proposed strategies to improve the state
of the practice, to address the identified pain points, and to improve software
quality. Combining their ideas with ours, we have the following list of
recommendations: increase documentation, increase testing by enriching datasets,
increase continuous integration usage, move to web applications, employ linters,
use peer reviews, design for change, add assurance cases, and incorporate a
``Generate All Things'' approach.

@@

\begin{table}[!ht]
\begin{center}
\begin{tabular}{ p{2.5cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1.2cm}p{1cm}p{0.8cm} }
\toprule
~ \ & \citet{USGS2019} & \citet{TobiasEtAl2018} & \citet{BrettEtAl2021} &
\citet{WilsonEtAl2016} & \citet{SmithAndRoscoe2018} & \citet{HerouxEtAl2008} &
\citet{ThielEtAl2020} & \citet{vanGompelEtAl2016} & \citet{OrvizEtAl2017} & MI\\
\midrule
LICENSE & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & &
\checkmark & \checkmark & \checkmark & C\\
README &  & \checkmark & \checkmark & \checkmark & \checkmark & & \checkmark &
\checkmark & \checkmark & C\\
CONTRIBUTING &  & \checkmark & \checkmark & \checkmark & \checkmark & &
\checkmark & \checkmark & \checkmark & R\\
CITATION &  &  &  & \checkmark & & & & \checkmark & \checkmark & U\\
CHANGELOG &  & \checkmark &  & \checkmark & \checkmark & & \checkmark &  &  & U\\
INSTALL &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark & U\\
\midrule
Uninstall &  &  &  &  & & & & \checkmark & &  \\
Dependency List &  &  & \checkmark & & \checkmark & & & \checkmark &  & R\\
Authors &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & U\\
Code of Conduct &  &  &  &  & & & \checkmark & & & R\\
Acknowledgements &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & U\\
Code Style Guide &  & \checkmark &  &  & & & \checkmark & \checkmark & \checkmark & R\\
Release Info. &  & \checkmark &  &  & & \checkmark & \checkmark & & & C\\
Prod.\ Roadmap &  &  &  &  & & \checkmark & \checkmark & \checkmark & & R\\
\midrule
Getting started &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark & R\\
User manual &  &  & \checkmark &  & & & \checkmark & & & C\\
Tutorials &  &  &  &  & & & \checkmark & & & U\\
FAQ &  &  &  &  & & & \checkmark & \checkmark & \checkmark & U\\
\midrule
Issue Track &  & \checkmark & \checkmark & & \checkmark & \checkmark &
\checkmark & & \checkmark & C\\
Version Control &  & \checkmark & \checkmark & \checkmark & \checkmark &
\checkmark & \checkmark & \checkmark & \checkmark & C\\ 
Build Scripts &  & \checkmark &  & \checkmark & \checkmark & \checkmark &
\checkmark & & \checkmark & U\\
\midrule
Requirements &  & \checkmark &  &  & & \checkmark &  &  & \checkmark & R\\
Design Doc.\ &  & \checkmark  & \checkmark &  & \checkmark & & \checkmark &
\checkmark& \checkmark & R\\
API Doc. &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark & R\\
Test Plan &  & \checkmark &  &  & & \checkmark & & & &  \\
Test Cases & \checkmark & \checkmark & \checkmark &  & \checkmark & \checkmark &
\checkmark & \checkmark & \checkmark & U\\
\bottomrule
\end{tabular}
\caption{Comparison of Recommended Artifacts in Software Development Guidelines
to Artifacts in MI Projects (C for Common, U for Uncommon and R for Rare)}
\label{Tbl_Guidelines}
\end{center}
\end{table}

@@

Although the MI community shows examples of 88\% (23 of 26) of the practices we
found in research software guidelines (Table~\ref{Tbl_Guidelines}), we did not
observe three recommended artifacts: i) Uninstall, ii) Test plans, and iii)
Requirements.  Uninstall is likely an omission caused by the focus on installing
software. Given the storage capacity of current hardware, developers and users
are not generally concerned with uninstall.  Moreover, as mentioned above,
uninstall is not particularly emphasized in existing recommendations.

Test plans describe the scope, approach, resources, and the schedule of planned
test activities \citep{VanVliet2000}.  The plan should cover details such as the
makeup of the testing team, automated testing tools and technology to employ,
the testing process, system tests, integration tests and unit tests. We did not
observe test plans for MI software, but that doesn't mean plans weren't created;
it means that the plans are not under version control. Test plans would have to
at least be implicitly created, since we observed test cases with reasonable
frequency for MI software (test cases are categorized as uncommon).

The other apparently neglected document is the requirements specification, which
records the functionalities, expected performance, goals, context, design
constraints, external interfaces and other quality attributes of the software
\citep{IEEE1998}.  If developers write requirements they typically are based on
a template, which provide documentation structure, guidelines, and rules.
Although there is no universally accepted template, examples include
\citet{ESA1991}, \citet{IEEE1998}, \citet{NASA1989}, and
\citet{RobertsonAndRobertson1999Vol}.

MI software is like other research software in its neglect of requirements
documentation.  Although requirements documentation is recommended by some
\citep{TobiasEtAl2018, HerouxEtAl2008, SmithAndKoothoor2016}, in practice
research software developers often do not produce a proper requirements
specification \citep{HeatonAndCarver2015}. \citet{SandersAndKelly2008}
interviewed 16 scientists from 10 disciplines and found that none of the
scientists created requirements specifications, unless regulations in their
field mandated such a document. \citet{Nguyen-HoanEtAl2010} showed requirements
are the least commonly produced type of documentation for research software in
general. When looking at the pain points for research software developers,
\citet{WieseEtAl2019} found that software requirements and management is the
software engineering discipline that most hurts scientific developers,
accounting for 23\% of the technical problems reported by study participants.
The lack of support for requirements is likely due to the perception that
up-front requirements are impossible for research software
\citep{CarverEtAl2007, SegalAndMorris2008}, but if we drop the insistence on
``up-front'' requirements, allowing instead for the requirements to be written
iteratively and incrementally, requirements are feasible \citep{Smith2016}.
\citet{SmithEtAl2007} provides a requirements template tailored to research
software.

Table~\ref{Tbl_Guidelines} shows several recommended artifacts that are rarely
observed in practice.  A theme among these rare artifacts is that, except for
the user-focused getting started manual, they are developer-focused.  The
dependency list, which is a list of software library dependencies, was rarely
observed, but this information is still likely present, just embedded in build
scripts. The other developer-focused and rare artifacts are as follows:

@@

@@

@@

As mentioned in Section~\ref{sec_result_correctness_verifiability}, we
identified five projects using CI/CD tools (about 17\% of the assessed
projects). We found which projects used CI/CD by examining the documentation and
source code of all projects. The count of CI/CD usage may actually be higher,
since traces of CI/CD usage may not always appear in a repository.  This was the
case for a study of LBM software, where interviews with developers showed that
more projects used CI/CD than was evident from repository artifacts alone
\citep{Michalski2021}.  The 17\% utilization for MI software contrasts with the
high frequency with which research software development guidelines recommend
continuous integration \citep{BrettEtAl2021, Brown2015, ThielEtAl2020,
Zadka2018, vanGompelEtAl2016}. Although there is currently little data available
on CI/CD utilization for research software, our impression is that CI/CD is not
yet common practice, despite its recommendation.  This is certainly the case for
LBM software, where usage numbers are similar to MI software, with only 12.5\%
of 24 LBM packages showing evidence of CI/CD in their repositories
\citep{Michalski2021}.  The survey of \citet{CarverEtAl2022} suggests higher use
of CI/CD with 54\% (54/100) of respondents indicating that they use it. However,
that survey measures something different from the current one by surveying
practitioners, rather than assessing projects.  Additional information on CI/CD
is given in the recommendations (Section~\ref{Sec_ContinuousIntegration}).

@@

\begin{itemize}
	\item Elimination of headaches associated with a separate integration phase
	\citep{Fowler2006}, \citep[p.\ 20]{HumbleAndFarley2010}. If developers
	postpone integration, integration problems are inevitable.  Continuous
	integration means that problems are immediately obvious and the source of
	the problem can be isolated to the small increment that was just committed.
	\item Detection and removal of bugs \citep{Fowler2006} via
	automated testing.  To improve productivity, defects are best discovered and
	fixed at the point where they are introduced \citep[p.\
	23]{HumbleAndFarley2010}.  Code is not the only source of errors; they are
	also found in the files and scripts related to configuration management
	\citep[p.\ 18]{HumbleAndFarley2010}.
	\item Everyone is always working on a stable base, since the rejection of
	inadequate commits means that the main branch will always be working.  A
	stable base will always pass all tests.  If the CI system uses generators
	and linters, it will also have current documentation and standard compliant
	code.  A stable base improves developer productivity, allowing them to focus
	on coding, testing, and documentation.
\end{itemize}

@@

\begin{itemize}
	\item A version control system \citep{Fowler2006}. To be effective, all
	files should be under version control, not just code files.  Anything that
	is needed to build, install and run the software should be under version
	control, including configuration files, build scripts, test harnesses, and
	operating system configuration files \citep[p. 19]{HumbleAndFarley2010}.
	Fortunately for the MI, as shown in Section~\ref{sec_score_maintainability}
	all our measured projects use version control.
	\item A fully automated build system \citep{Fowler2006}.  As \citet[p.\
	5]{HumbleAndFarley2010} point out, deploying software manually is an
	anti-pattern.  For MI software, Table~\ref{artifactspresent} shows 18 of 29
	packages (62\%) were observed to include build scripts.  Projects without a
	build system will need to add one to pursue using CI.
	\item An automated test system \citep{Fowler2006}. Building quality software
	involves creating automated tests at the unit, component, and acceptance
	test level, and executing these tests whenever someone makes a change to the
	code, its configuration, the environment, or the software stack that it runs
	on \citep[p.\ 83]{HumbleAndFarley2010}. As Table~\ref{artifactspresent}
	shows, test cases are in the uncommon category for MI software artifacts,
	which means that some MI projects will need to increase their testing
	automation if they wish to pursue CI.
	\item An automated system for other tasks, such as code checking,
	documentation building and web-site updating.  These other tasks are not
	essential to CI, but they can be incorporated to improve the quality
	of the code and the communication between developers and users. For
	instance, a static analysis (possibly via linters) of the code may find poor
	programming practice or lack of adherence to adopted coding standards.
	\item An integrated build system to pull everything together.  Every time
	there is a check-in (for instance a pull request), the integration server
	automatically checks out the sources onto the integration machine, starts a
	build, runs tests, and informs the committer of the results. 
\end{itemize}

@@

Continuous integration and delivery helps with addressing several pain points.
For instance, CI/CD helps reduce development time (\ppref{P_LackDevTime}) by
removing the need for a time-consuming integration stage and by automating
regression testing.  Automated regression tests also help with ensuring
correctness (\ppref{P_Correctness}) and the quality of reproducibility
(\qref{Q_Reproducibility}).

@@

\item \textbf{Modern technologies may improve frontend performance.} Web
applications with only a frontend usually perform worse than native
applications. However, new technologies may ease this difference. For example,
some JavaScript libraries can help the frontend harness the power of the
computer's GPU and accelerate graphical computing. In addition, there are new
frameworks helping developers with cross-platform compatibility. For example,
the \href{https://flutter.dev/}{Flutter} project enables support for web,
mobile, and desktop OS with one codebase.  Other options include
\href{https://vuejs.org/} {Vue}, \href{https://angular.io/} {Angular} and
\href{https://reactjs.org/} {React}, and \href{https://elm-lang.org/}{Elm}.

@@

A linter is a tool that statically analyzes code to find programming errors,
suspicious constructs, and stylistic inconsistencies \citep{Wikipedia2022_Lint}.
Linters can be used as an ad hoc check for code files, but they really come into
their own when used as part of a CI system, as discussed in
Section~\ref{Sec_ContinuousIntegration}. Almost none of the research software
guidelines that we consulted, summarized in Section~\ref{Sec_CompareArtifacts},
mention linters.  The one exception is \citet{ThielEtAl2020}.  Despite the lack
of mention in the guidelines, we believe that linters have the potential to
improve code quality at a relatively low cost.

@@

\citet{Fagan1976} began work on rigorous review via code inspection.  Elements
of a typical inspection include reviewing the code against a checklist (checking
the consistency of variable names, look for terminating loops, etc.), performing
specific review tasks (such as summarizing the code's purpose, cross-referencing
the code to the technical manual, creating a data dictionary for a given module,
etc.) Rigorous inspection finds 60-65\% of latent defects on average, and often
tops 85\% in defect removal efficiency \citep{Jones2008}. The success rate of
code inspection is generally higher than most forms of testing, which average
between 30 --- 35\% for defect removal efficiency \citep{EbertAndJones2009,
Jones2008}. For research software, \citep{KellyAndShepard2000} show a task based
inspection approach can be effective. Task based inspection is an ideal fit with
an issue tracking system, like GitHub.  The review tasks can be issues, so that
they can be easily assigned, monitored and recorded. Potential issues include
assigning junior developers to test getting-started tutorial and installation
instructions.

As indicated in Section~\ref{Sec_CompareMethodologies} some MI projects use
modern code review, via issue tracking and the use of GitHub.  Those MI projects
not incorporating modern code review would likely benefit by adopting it.
Although a rigorous code inspection is likely not worth the required resources,
for critical parts of the code, developers may want to adopt a more rigorous
approach. For instance, developers may drop the modern trend of
asynchronous review and instead occasionally use synchronous review to help
uncover errors and disseminate best practices throughout the team.  For
instance, teams could periodically meet, either in-person or virtually, and have
junior members walk through their code.  In-person reviews will likely help
realize the benefits of modern code review noticed by
\citet{BirdAndBacchelli2013}: defect detection, knowledge transfer, increased
team awareness, and creation of alternative solutions to problems.

Due to improving code quality and increasing knowledge transfer, peer review
addresses the same pain points and qualities as linters
(Section~\ref{Sec_Linters}): \ppref{P_LackDevTime}, \ppref{P_TechnologyHurdles},
and \qref{Q_Maintainability}. Peer review can potentially find misunderstandings
in how the code implements the required theory, which will improve the
software's correctness (\ppref{P_Correctness}). The benefits of peer review for
addressing pain points can be increased by extending the review from just code,
to also reviewing all software artifacts, including documentation, build
scripts, test cases and the development process itself.

@@

In our ``state of the practice'' assessment exercise for LBM software
\citep{SmithEtAl2022}, we noticed that LBM developers implicitly used
modularization based on the principle of design for change to improve
maintainability (\qref{Q_Maintainability}).  We recommend that MI developers use
the same principle for their modularizations.  Although the advice to modularize
research software to handle complexity is common \citep{WilsonEtAl2014,
StewartEtAl2017, Storer2017}, specific guidelines on how to divide the software
into modules is less prevalent.  Not every decomposition is a good design for
supporting change, as shown by \citet{Parnas1972a}.  For instance, a design with
low cohesion and high coupling \citep[p.\ 48]{GhezziEtAl2003} will make change
difficult. Especially in research software, where change is inevitable,
designers need to produce a modularization that supports change.
\citet{JungEtAl2022} points out that ocean modelling software is currently
feeling the pain of not emphasizing modularization in legacy code.

@@

As the LBM examples above illustrate, developers can accomplish design for
change by first identifying likely changes, either implicitly or explicitly, and
second by hiding each likely change behind a well-defined module interface. This
approach mirrors the recommendations from \citet{Parnas1972a}.
Section~\ref{Sec_CompareArtifacts} lists ideas for how to document the design,
including the likely changes, so that they are more visible to others.

@@

\citet{RunesonAndHost2009} defines construct validity as the adopted
metrics representing what they are intended to measure. Our construct threats are
often related to how we assume our measurements influences the various software
qualities, as summarized in Section~\ref{sec_grading_software}. Specifically,
our construct validity related threats include the following:
